What paradigms can webcam eye-tracking be used for? Attempted replications of 5 ``classic'' cognitive science experiments

### 2023-10-05 manuscript.tex
re-knit pdf

Finally, several papers have used the \texttt{WebGazer.js} library (via different interfaces) to replicate visual world paradigm studies. Slim and Hartsuiker (2022) used the \texttt{PCIbex} online experiment platform (Zehr \& Schwarz, 2018) to replicate a study in which participants (n=90) listened to sentences (e.g., Mary reads a letter) while viewing four pictures displayed across the four quadrants of the screen (e.g., a letter, a backpack, a car and a wheelchair) (Dijkgraaf, Hartsuiker, \& Duyck, 2017). When the verb in the sentence was \emph{constraining} (e.g., read) with respect to the display (only the letter would be an appropriate continuation), listeners made more fixations to the target image (letter) than when the verb was \emph{neutral} (e.g., Mary steals a letter), as in previous work. However, they observed a substantial delay of \textasciitilde200ms in the onset of the effect relative to the in-lab study with an infrared eyetracker and a reduction in effect size (despite a threefold increase in sample size relative to the original). The authors noted that delay was particularly problematic given that the purpose of the original study was to capture aspects of predictive or anticipatory processing: in the original study, but not in the web replication, the difference between \emph{constraining} and \emph{neutral} conditions emerged before the onset of the final noun (e.g., letter). Similarly, Degen, Kursat, and Leigh (2021) reported that effects of scalar implicature processing (comparing looks to the target in a four quadrant display for sentences such as ``Click on the girl that has \emph{some} apples'' vs.~``Click on the girl that has \emph{three} apples'' ) were smaller and delayed relative to those observed in the lab (Sun \& Breheny, 2020).

In contrast, using the \texttt{Gorilla} experiment platform (Anwyl-Irvine et al., 2020), Prystauka, Altmann, and Rothman (2023) observed robust effects of verb semantics and lexical interference in similar time windows to previous in-lab studies (Altmann \& Kamide, 1999; Kukona, Cho, Magnuson, \& Tabor, 2014), though a direct comparison was not possible because their studies used different materials than the in-lab experiments they were conceptually replicating. Furthermore, using \texttt{jsPsych}, Vos, Minor, and Ramchand (2022) closely replicated the magnitude and timecourse of effects in a lab-based visual world paradigm where the regions of interest (ROIs) consisted of the left and right halves of the display, suggesting that the concerns about poor temporal resolution can be mitigated. Indeed, the eye-tracking plug-in in \texttt{jsPsych} uses a fork of \texttt{Webgazer.js} in which certain modifications have been made to minimize processing time. \footnote{See discussion at \url{https://github.com/jspsych/jsPsych/discussions/1892}}

In sum, webcam eye-tracking has been used to varying degrees of success across a small set of paradigms that are used in cognitive science research. One configuration, \texttt{jsPsych} with a modification of \texttt{WebGazer.js}, appears to have circumvented initial limitations in temporal precision (Krajbich et al., 2010; Vos et al., 2022) but has only been tested with paradigms with minimal requirements in terms of spatial precision (both experiments used the two sides of the display as the ROIs). In the current work, we evaluate a broad variety of paradigms in terms of their suitability for webcam eye-tracking, in order to provide guidance for cognitive scientists aiming to increase and diversify participation in their eye-tracking-based research.

@@

In order to validate the online eyetracking methodology, with the particular configuration known to have the greatest temporal precision, \texttt{jsPsych} and a modification of \texttt{Webgazer}, we set out to reproduce five previously published studies representing a variety of questions, topics, and paradigms. The goal was to examine the strengths and weaknesses of webcam eye-tracking for common paradigms in cognitive science. Ideally, we would only attempt to replicate studies where the original measurements have small standard errors and are known to replicate; otherwise, it can be difficult to distinguish a failure of the method (online eyetracking does not work) from a failure of the original study to replicate.

In practice, replications (successful or otherwise) have only been reported for a small number of studies, so we ultimately included some studies with unknown replicability. We addressed this in several ways. First, replicating five very different studies from different research traditions decreases our reliance on any one study. Second, we include several `'sanity check'' analyses, such as the correlation between calibration accuracy and effect size. (If the effect is real but there is noise from low-accuracy eyetracking, this correlation should be substantial.) Third, for two of the studies, we had comparison data collected in-lab either using \texttt{jsPsych} or a more traditional eyetracker technology, allowing us to direct assess the impact of differences in subject population and equipment.

@@

@@

@@

The visual stimuli were created through Canva and depicted an agent
accompanied by four to five objects in the scene (see Figure ~\ref{fig:E1-example-trial}). On
critical trials, participants heard one of two sentences associated with
the scene. In the restrictive condition, the sentence (e.g., ``The boy
will eat the cake'') contained a verb (e.g., ``eat'') which restricts the
set of possible subsequent referents (e.g., to edible things). Only the
target object (e.g., the cake) was semantically consistent with the
verb's meaning. In the non-restrictive condition, the sentence (e.g.,
``The boy will move the cake'') contained a verb (e.g., ``move'') which does
not restrict the set of possible subsequent referents. The target object
(e.g., the cake) as well as the distractor objects (e.g., the train, the
ball, etc.) were semantically consistent with the verb's meaning. Both
sentences were compatible with the scene, such that the correct keyboard
response for the critical trials was ``yes.'' Filler trials consisted of
scenes that looked similar to critical scenes but were paired with
inappropriate sentences. The correct keyboard response for the filler
trials was ``no.''

@@

The experimental design is visually depicted in Figure~\ref{fig:E3-design-schematic}

@@

Images were selected XXX\ldots{}

There were two modifications we made to the methods of the original
experiment. As we are only replicating the declarative memory component
of the original experiment, we did not have a ``priming group.''
Therefore, we followed only the procedure for the ``looking group.''
Additionally, for each section of the study, the stimuli was presented
on a single screen instead of two screens due to the constraints of the
online experiment format.

@@

@@

The stimuli, experimental code, and data and analysis scripts can be
found on the Open Science Framework at the following link,
\url{https://osf.io/x3c49/}. The pre-registration
for the study can be found at \url{https://osf.io/3v4pg}.

@@

@@

All stimuli, experiment scripts, data, and analysis scripts are
available on the Open Science Framework at \url{https://osf.io/eubsc/}
(\url{https://osf.io/eubsc/}). The study pre-registration is available at
\url{https://osf.io/tv57s} (\url{https://osf.io/tv57s}).

@@

During each trial of the main task, two faces were displayed on the two
halves of the screen, one on the left and one on the right (as in Figure~\ref{fig:E5-example-trial}). Participants were randomly assigned to one of two tasks:
attractiveness or shape judgment. In the attractiveness task,
participants were asked to chose the more attractice face in the pair
and in the shape judgment task participants were asked to pick the face
that appeared rounder. They pressed the ``a'' key on their keyboard to
select the face on the left and the ``d'' key to select the face on the
right. A fixation cross appeared in the center of the screen between
each set of faces. Participants were asked to look at this fixation
cross in order to reset their gaze in between trials (???). The order of
the 19 face pairs was random for each participant.

@@

The faces in our replication were selected from a set of 1,000 faces
within the Flickr-Faces-HQ Dataset. (The face images used in Shimojo et
al.~were from the Ekman face database and the AR face database.) These
images were chosen because the person in each image was looking at the
camera with a fairly neutral facial expression and appeared to be over
the age of 18. 27 participants were recruited on Prolific to participate
in stimulus norming (for attractiveness). They were paid \$XX for
completing the experiment. Data from 3 participants was excluded because
their mode response made up more than 50\% of their total responses
, for a total of 24 participants
in the norming. They each viewed all 172 faces and were asked to rate
them on a scale from 1 (less attractive) to 7 (more attractive) using a
slider. Faces were presented one at a time and in a random order for
each participant. Following Shimojo et al., 19 face pairs were made by
matching two faces that had a difference in mean attractiveness ratings
that was 0.25 points or lower and that matched in gender, race, and age
group (young adult, adult, or older adult).

@@

In the original study, a video-based eye tracker was used. The eye
movements of participants were recorded with a digital camera
downsampled to 33.3 Hz, with eye position was then determined
automatically with MediaAnalyzer software. In our study, subjects
supplied their own cameras, so hardware sampling rate varied. However,
data was collected at 20 Hz.{[}TODO - CONFIRM{]}

@@

\hypertarget{combined-analyses}{%
\section{Combined Analyses?}\label{combined-analyses}}

\begin{itemize}
\tightlist
\item
  Pooling data from all experiments we can look at patterns in the
  calibration and validation data
\end{itemize}

@@

\begin{itemize}
\item
  E1:
\item
  E2:

  \begin{itemize}
  \tightlist
  \item
    replication of key result with 4 quadrants
  \item
    calib quality doesn't seem to matter
  \item
    if attempting to control where people are looking, think about modifying task\ldots{}
  \end{itemize}
\item
  E3:
\item
  E4:
\item
  E5:
\end{itemize}

@@

@@

>>>>

Several papers have used the \texttt{WebGazer.js} library (via different interfaces) to replicate visual world paradigm studies. Slim and Hartsuiker (2022) used the \texttt{PCIbex} online experiment platform (Zehr \& Schwarz, 2018) to replicate a study in which participants (n=90) listened to sentences (e.g., Mary reads a letter) while viewing four pictures displayed across the four quadrants of the screen (e.g., a letter, a backpack, a car and a wheelchair) (Dijkgraaf, Hartsuiker, \& Duyck, 2017). When the verb in the sentence was \emph{constraining} (e.g., read) with respect to the display (only the letter would be an appropriate continuation), listeners made more fixations to the target image (letter) than when the verb was \emph{neutral} (e.g., Mary steals a letter), as in previous work. However, they observed a substantial delay of \textasciitilde200ms in the onset of the effect relative to the in-lab study with an infrared eyetracker and a reduction in effect size (despite a threefold increase in sample size relative to the original). The authors noted that delay was particularly problematic given that the purpose of the original study was to capture aspects of predictive or anticipatory processing: in the original study, but not in the web replication, the difference between \emph{constraining} and \emph{neutral} conditions emerged before the onset of the final noun (e.g., letter). Similarly, Degen, Kursat, and Leigh (2021) reported that effects of scalar implicature processing (comparing looks to the target in a four quadrant display for sentences such as ``Click on the girl that has \emph{some} apples'' vs.~``Click on the girl that has \emph{three} apples'' ) were smaller and delayed relative to those observed in the lab (Sun \& Breheny, 2020).

In contrast, using the \texttt{Gorilla} experiment platform (Anwyl-Irvine et al., 2020), Prystauka, Altmann, and Rothman (2023) observed robust effects of verb semantics and lexical interference in similar time windows to previous in-lab studies (Altmann \& Kamide, 1999; Kukona, Cho, Magnuson, \& Tabor, 2014), though a direct comparison was not possible because their studies used different materials than the in-lab experiments they were conceptually replicating. Furthermore, using \texttt{jsPsych}, Vos, Minor, and Ramchand (2022) closely replicated the magnitude and timecourse of effects in a lab-based visual world paradigm where the regions of interest (ROIs) consisted of the left and right halves of the display, suggesting that the concerns about poor temporal resolution can be mitigated. Indeed, the eye-tracking plug-in in \texttt{jsPsych} uses a fork of \texttt{Webgazer.js} in which certain modifications have been made to minimize processing time. \footnote{See discussion at \url{https://github.com/jspsych/jsPsych/discussions/1892}}
Finally, Van der Cruyssen et al. (2023) used \texttt{WebGazer.js} in conjunction with \texttt{PsychoJS} to conceptually replicate three paradigms, one which is analogous to Yang and Krajbich (2021), another testing the novelty preference (tendency for learners to look more at an object that they have not yet studied), and a visual world paradigm study with 4 objects, similar to Prystauka et al. (2023). These replications were successful but they noted reductions in effect sizes relative to an in-lab dataset. (They did not address whether there were temporal delays.)

In sum, webcam eye-tracking has been used to varying degrees of success across a small set of paradigms that are used in cognitive science research. One configuration, \texttt{jsPsych} with a modification of \texttt{WebGazer.js}, appears to have circumvented initial limitations in temporal precision (Krajbich et al., 2010; Vos et al., 2022) but has only been tested with paradigms with minimal requirements in terms of spatial precision (both experiments used the two sides of the display as the ROIs). Further, the extent to which noise in the measurement is due to the software and hardware limitations as opposed to the difference between the in-lab and online settings remains unclear. In the current work, we evaluate a broad variety of paradigms in terms of their suitability for webcam eye-tracking in order to provide guidance for cognitive scientists aiming to increase and diversify participation in their eye-tracking-based research.

@@

In order to validate the online eyetracking methodology, with the particular configuration known to have the greatest temporal precision, \texttt{jsPsych} and a modification of \texttt{Webgazer}, we set out to reproduce five previously published studies representing a variety of questions, topics, and paradigms. The goal was to examine the strengths and weaknesses of webcam eye-tracking for common paradigms in cognitive science, across a broad range of research areas.

@@

Studies with large effect sizes and which are known to replicate are ideal targets for further replication; otherwise, it can be difficult to distinguish a failure of the method from a failure of the original study to replicate. In practice, replications (successful or otherwise) have only been reported for a small number of studies, so we ultimately included some studies with unknown replicability. We addressed this in several ways. First, replicating five very different studies from different research traditions decreases our reliance on any one study. Second, we include several `'sanity check'' analyses, such as the correlation between calibration accuracy and effect size. (If the effect is real but there is noise from low-accuracy eyetracking, this correlation should be substantial.) Third, for two of the studies, we had comparison data collected in-lab either using \texttt{jsPsych} or a more traditional eyetracker technology, allowing us to directly assess the impact of differences in subject population and equipment/setting.

@@

\hypertarget{pre-registration}{%
\subsection{Pre-registration}\label{pre-registration}}

These data were collected within the context of an undergraduate research methods course. Groups of students (co-authors) designed and programmed experiments in jsPsych, pre-registered their planned analyses, and collected data through Prolific under the supervision of the first author. The OSF repositories associated with these experiments are linked in the methods sections of each individual study. Note that in the current paper we expand on those pre-registered analyses (e.g., including analyses of the calibration quality). All analysis code underlying this paper can be found in the Github repository: \url{https://github.com/jodeleeuw/219-2021-eyetracking-analysis}

@@

The visual stimuli were created through Canva and depicted an agent
accompanied by four to five objects in the scene (see Figure~\ref{fig:E1-example-trial}). On
critical trials, participants heard one of two sentences associated with
the scene. In the restrictive condition, the sentence (e.g., ``The boy
will eat the cake'') contained a verb (e.g., ``eat'') which restricts the
set of possible subsequent referents (e.g., to edible things). Only the
target object (e.g., the cake) was semantically consistent with the
verb's meaning. In the non-restrictive condition, the sentence (e.g.,
``The boy will move the cake'') contained a verb (e.g., ``move'') which does
not restrict the set of possible subsequent referents. The target object
(e.g., the cake) as well as the distractor objects (e.g., the train, the
ball, etc.) were semantically consistent with the verb's meaning. Both
sentences were compatible with the scene, such that the correct keyboard
response for the critical trials was ``yes.'' Filler trials consisted of
scenes that looked similar to critical scenes but were paired with
inappropriate sentences. The correct keyboard response for the filler
trials was ``no.''

@@

The experimental design is visually depicted in Figure~\ref{fig:E3-design-schematic}.

There were two modifications we made to the methods of the original
experiment. As we were only replicating the declarative memory component
of the original experiment, we did not have a ``priming group.''
Therefore, we followed only the procedure for the ``looking group.''
Additionally, for each section of the study, the stimuli were presented
on a single screen instead of two screens due to the constraints of the
online experiment format.

@@

\textcolor{red}{Images were selected XXX...}

@@

As in Manns et al. (2000), participants looked more at novel images than previously seen images. This effect was consistent for ROIs based on the images and for the coarser ROIs based on two halves of the display. A day later, participants were also able to discriminate the images they had seen from foil images they had not seen during the previous session. However, there was no evidence that memory performance on day 2 was related to looking time on day 1. Calibration quality did not appear to impact this relationship.

@@

The stimuli, experimental code, and data and original analysis scripts can be
found on the Open Science Framework at the following link,
\url{https://osf.io/x3c49/}. The pre-registration
for the study can be found at \url{https://osf.io/3v4pg}.

@@

As in Ryskin et al. (2017) and Snedeker and Trueswell (2004), listeners' gaze patterns during sentences with globally ambiguous syntactic interpretations differed depending on the bias of the verb (i.e., modifier, instrument or equi). For modifier-biased verbs, participants looked more quickly at the target animal and less at the potential instrument than for instrument-biased verbs (and equi-biased verbs elicited a gaze pattern between these extremes). This pattern was stronger for those who achieved higher calibration accuracy and when quadrant-based ROIs were used compared to image-based ROIs.

@@

All stimuli, experiment scripts, data, and analysis scripts are
available on the Open Science Framework at \url{https://osf.io/eubsc/}.
The study pre-registration is available at
\url{https://osf.io/tv57s}.

@@

During each trial of the main task, two faces were displayed on the two
halves of the screen, one on the left and one on the right (as in Figure~\ref{fig:E5-example-trial}). Participants were randomly assigned to one of two tasks:
attractiveness or shape judgment. In the attractiveness task,
participants were asked to chose the more attractice face in the pair
and in the shape judgment task participants were asked to pick the face
that appeared rounder. They pressed the ``a'' key on their keyboard to
select the face on the left and the ``d'' key to select the face on the
right. A fixation cross appeared in the center of the screen between
each set of faces. Participants were asked to look at this fixation
cross in order to reset their gaze in between trials. The order of
the 19 face pairs was random for each participant.

@@

The faces in our replication were selected from a set of 1,000 faces
within the Flickr-Faces-HQ Dataset. (The face images used in Shimojo et
al.~were from the Ekman face database and the AR face database.) These
images were chosen because the person in each image was looking at the
camera with a fairly neutral facial expression and appeared to be over
the age of 18. 27 participants were recruited on Prolific to participate
in stimulus norming (for attractiveness). They each viewed all 172 faces and were asked to rate
them on a scale from 1 (less attractive) to 7 (more attractive) using a
slider. Faces were presented one at a time and in a random order for
each participant. Data from 3 participants were excluded because
their mode response made up more than 50\% of their total responses, for a total of 24 participants
in the norming.

Following Shimojo et al., 19 face pairs were selected by
identifying two faces that 1) had a difference in mean attractiveness ratings
that was 0.25 points or lower and 2) matched in gender, race, and age
group (young adult, adult, or older adult).

@@

In the original study, a video-based eye tracker was used. The eye
movements of participants were recorded with a digital camera
downsampled to 33.3 Hz, with eye position was then determined
automatically with MediaAnalyzer software. In our study, subjects
supplied their own cameras, so hardware sampling rate varied. However,
data was collected at 20 Hz.\textcolor{red}{[TODO - CONFIRM]}

@@

Qualitatively, the results are similar to those of Shimojo et al., such that participants look more at the face that they ultimately choose. This gaze bias appears to be stronger for decisions about face attractiveness than shape, though this is not supported by the statistical analysis approach used in the original paper. The gaze patterns remained consistent for participants with better calibration accuracy.

@@

We conducted 5 attempted replication studies using different experimental paradigms from across the cognitive sciences. All were successfully implemented in \texttt{jsPsych} using the \texttt{webgazer} plugin, but replication success was mixed. Experiment 1 had the smallest ROIs due to the use of an integrated visual scene with 5 to 6 ROIs of varying size per scene, as opposed to ROIs corresponding to display halves or quadrants. Both attempts to replicate Altmann and Kamide (1999) were unsuccessful, despite the success of previous in-lab replications using infrared eye-tracking (James, Minnihan, \& Watson, 2023). A previous conceptual replication of this paradigm using webcam-based eye-tracking {[}prystauka2023online{]} was successful but used a 4-quadrant visual world paradigm, rather than the ``naturalistic'' scenes used in the original study and in the current replication attempts. It is worth noting that removing variability related to participant environments (by conducting the webcam-tracking study in the lab) did not appear to improve the sensitivity of the paradigm. The primary limitation is likely to be the size of the ROIs.

Experiment 2 used the 4 quadrants of the participant's screen as ROIs. As in Johansson and Johansson (2014) and Spivey and Geng (2001), participants spontaneously looked to blank ROIs which previously contained to-be-remembered pictures. These results appeared to be robust to calibration quality. An additional manipulation, instructing participants to keep gaze fixed on a central point, was not successful. One possibility is that participants are less motivated to follow such instructions when an experimenter is not present in the same room with them. It may be possible to improve performance by emphasizing that this is an important aspect of the experiment or by providing additional training/practice in keeping the eyes still on one particular point.

Experiment 3 used 2 large ROIs (halves of the display in one analysis) and successfully replicated the novelty preference in terms of gaze duration shown in (\textbf{mannsVisualPairedcomparisonTask200?}). However, the subtler relationship between gaze duration and recognition memory on day 2 was not replicated, despite the fact that participants were able to discriminate pictures they had seen from those they hadn't seen during that delayed test. Calibration quality did not appear to impact this relationship. More work is needed to understand whether delay manipulations can be practically combined with webcam eye-tracking.

Experiment 4 used the 4 quadrants of the participant's screen as ROIs. As in Ryskin et al. (2017), listeners used knowledge of the co-occurrence statistics of verbs and syntactic structures to resolve ambiguous linguistic input (``Rub the frog with the mitten''). Across multiple time windows, participants looked more at potential instruments (mitten), when the verb (rub) was one that was more likely to be followed by a prepositional phrase describing an instrument with which to perform the action , as opposed to describing the recipient of the action (frog). Despite the qualitative replication of past findings, the overall rates of looks to various objects were much lower than in an in-lab study using infrared eye-tracking. This reduction may be related to measurement quality: effect sizes were greater for participants with higher calibration accuracy. Using the full quadrants as ROIs, rather than bounding boxes around the four images, also appeared to improve the measurement of the effect. Crucially, there was no evidence of a delay in the onset of effects relative to in-lab work, indicating that the modifications to \texttt{webgazer} that are made within the \texttt{jsPsych} plug-in successfully address the issues noted by Dijkgraaf et al. (2017) and suggesting that this methodology can be fruitfully used to investigate research questions related to the timecourse of processing.

Experiment 5, similar to Experiment 3, used 2 large ROIs (or halves of the display). As in Shimojo et al. (2003) and in the recent webcam-based replication by Yang and Krajbich (2021), participants (qualitatively) look more at the face that they ultimately choose. This gaze bias appears to be stronger for decisions about face attractiveness than shape, though this is not supported by the statistical analysis approach used in the original paper. The gaze patterns remained consistent for participants with better calibration accuracy.

In sum, the \texttt{webgazer} plug-in for \texttt{jsPsych} can be fruitfully used to conduct a variety of cognitive science experiments on the web, provided the limitations of the methodology are carefully considered. Studies with ROIs that take up half or a quarter of the participant's display, which encompasses a large number of common paradigms, are very likely to be successful, even when testing questions related to the timecourse of processing. However, the smaller the ROIs, the more important the calibration becomes. For instance, studies with 4 ROIs may want to exclude data from participants with less than 75\% validation accuracy, whereas studies using 2 halves of the display as ROIs may not need to be so conservative. Studies with smaller ROIs (see Experiment 1) may not be appropriate for webcam eye-tracking in its current form.

@@

\leavevmode\vadjust pre{\hypertarget{ref-james2023language}{}}%
James, A. N., Minnihan, C. J., \& Watson, D. G. (2023). Language experience predicts eye movements during online auditory comprehension. \emph{Journal of Cognition}, \emph{6}(1).

@@

\leavevmode\vadjust pre{\hypertarget{ref-van2023validation}{}}%
Van der Cruyssen, I., Ben-Shakhar, G., Pertzov, Y., Guy, N., Cabooter, Q., Gunschera, L. J., \& Verschuere, B. (2023). The validation of online webcam-based eye-tracking: The replication of the cascade effect, the novelty preference, and the visual world paradigm. \emph{Behavior Research Methods}, 1--14.
